---
title: "Analysis of Streaming Transactions"
author: "Kanad Motiwale"
term: "Fall 2025"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
    toc_depth: 2
    number_sections: true
    theme: paper
    highlight: tango
  pdf_document:
    toc: true
    toc_depth: '2'
subtitle: Based on Data Collected by SportsTV
---


```{r setup, include=FALSE}
# Setup chunk - load libraries and connect to database
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Function to install packages if not already installed
install_if_missing <- function(packages) {
  for (pkg in packages) {
    if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
      install.packages(pkg, repos = "https://cran.rstudio.com/")
      library(pkg, character.only = TRUE)
    }
  }
}

# Install and load required packages
required_packages <- c("RMySQL", "DBI", "knitr", "ggplot2")
install_if_missing(required_packages)

# Load libraries
library(RMySQL)
library(DBI)
library(knitr)
library(ggplot2)

# Load database configuration
source("config.R")

# Database connection
mysqlConn <- dbConnect(
  RMySQL::MySQL(),
  host = DB_HOST,
  port = DB_PORT,
  dbname = DB_NAME,
  user = DB_USER,
  password = DB_PASSWORD
)
```

# Executive Summary
```{r executiveSummary, include=FALSE}
# Get summary statistics using JOINs
min_date <- dbGetQuery(mysqlConn, "
  SELECT MIN(dd.full_date) 
  FROM fact_streaming f
  JOIN dim_date dd ON f.date_id = dd.date_id
")[[1]]

max_date <- dbGetQuery(mysqlConn, "
  SELECT MAX(dd.full_date) 
  FROM fact_streaming f
  JOIN dim_date dd ON f.date_id = dd.date_id
")[[1]]

total_streams <- dbGetQuery(mysqlConn, "SELECT SUM(total_streams) FROM fact_streaming")[[1]]

total_countries <- dbGetQuery(mysqlConn, "
  SELECT COUNT(DISTINCT country_id) FROM fact_streaming
")[[1]]

total_sports <- dbGetQuery(mysqlConn, "
  SELECT COUNT(DISTINCT sport_id) FROM fact_streaming
")[[1]]

# Format for display
min_date_formatted <- format(as.Date(min_date), "%B %d, %Y")
max_date_formatted <- format(as.Date(max_date), "%B %d, %Y")
total_streams_formatted <- format(total_streams, big.mark = ",")
```

This report analyzes streaming transactions for SportsTV Germany, covering `r min_date_formatted` through `r max_date_formatted`. The analysis includes **`r total_streams_formatted`** total streaming events across **`r total_countries`** countries featuring **`r total_sports`** sports.

**Data Quality Handling:** The ETL process filtered out transactions with NULL values in critical fields (date, sport, country, or duration). From 1.18 million raw records, 996,091 valid transactions were loaded, representing high-quality data suitable for analysis. All fact table records maintain referential integrity with zero NULL values.

---

# Growth of Streaming Transactions Over Time

## Analysis by Sport

This section examines the growth trends in streaming activity across different sports over all available years.
```{r growthBySport}
# Query with JOINs to dimension tables
sql <- "
SELECT 
    dd.year,
    ds.sport_name as sport,
    SUM(f.total_streams) as total_streaming_events,
    SUM(f.total_minutes_streamed) as total_streaming_minutes
FROM fact_streaming f
INNER JOIN dim_date dd ON f.date_id = dd.date_id
INNER JOIN dim_sport ds ON f.sport_id = ds.sport_id
GROUP BY dd.year, ds.sport_name
ORDER BY dd.year, ds.sport_name
"

df_growth <- dbGetQuery(mysqlConn, sql)

# Display table
kable(df_growth, 
      col.names = c("Year", "Sport", "Total Streaming Events", "Total Streaming Minutes"),
      format.args = list(big.mark = ","),
      caption = "Table 1: Streaming Activity by Year and Sport")
```
```{r growthChart, fig.width=10, fig.height=6}
# Create line chart for streaming events
ggplot(df_growth, aes(x = year, y = total_streaming_events, color = sport, group = sport)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "Growth of Streaming Events by Sport",
    subtitle = "Total number of streaming events per year",
    x = "Year",
    y = "Number of Streaming Events",
    color = "Sport"
  ) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )
```
```{r growthChartMinutes, fig.width=10, fig.height=6}
# Create line chart for streaming minutes
ggplot(df_growth, aes(x = year, y = total_streaming_minutes, color = sport, group = sport)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "Growth of Streaming Time by Sport",
    subtitle = "Total minutes streamed per year",
    x = "Year",
    y = "Total Streaming Minutes",
    color = "Sport"
  ) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )
```

### Key Findings
```{r growthInsights}
# Calculate year-over-year growth
most_recent_year <- max(df_growth$year)
previous_year <- most_recent_year - 1

current_total <- sum(df_growth[df_growth$year == most_recent_year, "total_streaming_events"])
previous_total <- sum(df_growth[df_growth$year == previous_year, "total_streaming_events"])

growth_pct <- round(((current_total - previous_total) / previous_total) * 100, 1)

# Most popular sport
sport_totals <- aggregate(total_streaming_events ~ sport, data = df_growth, sum)
most_popular_sport <- sport_totals[which.max(sport_totals$total_streaming_events), "sport"]
```

- Streaming events grew by **`r growth_pct`%** from `r previous_year` to `r most_recent_year`
- **`r most_popular_sport`** is the most streamed sport with **`r format(max(sport_totals$total_streaming_events), big.mark=",")`** total streaming events across all years

---

# Weekly Streaming Patterns

## Most Recent Year Analysis

This section examines weekly streaming patterns for the most recent year of data.
```{r weeklyPatterns}
# Get most recent year
recent_year <- dbGetQuery(mysqlConn, "
  SELECT MAX(dd.year) as max_year 
  FROM fact_streaming f
  JOIN dim_date dd ON f.date_id = dd.date_id
")$max_year

# Query with JOIN
sql <- paste0("
SELECT 
    dd.week,
    SUM(f.total_streams) as streaming_events
FROM fact_streaming f
INNER JOIN dim_date dd ON f.date_id = dd.date_id
WHERE dd.year = ", recent_year, "
GROUP BY dd.week
ORDER BY dd.week
")

df_weekly <- dbGetQuery(mysqlConn, sql)

# Display table
kable(df_weekly, 
      col.names = c("Week Number", "Total Streaming Events"),
      format.args = list(big.mark = ","),
      caption = paste0("Table 2: Weekly Streaming Events in ", recent_year))
```
```{r weeklyChart, fig.width=10, fig.height=6}
# Create bar chart
ggplot(df_weekly, aes(x = week, y = streaming_events)) +
  geom_col(fill = "steelblue") +
  labs(
    title = paste0("Weekly Streaming Events in ", recent_year),
    subtitle = "Number of streaming events per week",
    x = "Week Number",
    y = "Streaming Events"
  ) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12)
  )
```

### Weekly Statistics
```{r weeklyStats}
avg_weekly <- round(mean(df_weekly$streaming_events), 0)
max_week <- df_weekly[which.max(df_weekly$streaming_events), "week"]
max_events <- max(df_weekly$streaming_events)
min_week <- df_weekly[which.min(df_weekly$streaming_events), "week"]
min_events <- min(df_weekly$streaming_events)
```

- Average weekly streaming events: **`r format(avg_weekly, big.mark=",")`**
- Peak week: Week **`r max_week`** with **`r format(max_events, big.mark=",")`** events
- Lowest week: Week **`r min_week`** with **`r format(min_events, big.mark=",")`** events

---

# Streaming Analytics by Sport and Country

## Average Streaming Time and Total Streams
```{r sportCountryAnalysis}
# Query with JOINs
sql <- "
SELECT 
    ds.sport_name as sport,
    dc.country_name as country,
    SUM(f.total_streams) as total_streams,
    ROUND(AVG(f.avg_minutes_per_stream), 2) as avg_streaming_minutes
FROM fact_streaming f
INNER JOIN dim_sport ds ON f.sport_id = ds.sport_id
INNER JOIN dim_country dc ON f.country_id = dc.country_id
GROUP BY ds.sport_name, dc.country_name
ORDER BY ds.sport_name, dc.country_name
"

df_sport_country <- dbGetQuery(mysqlConn, sql)

# Display table
kable(df_sport_country, 
      col.names = c("Sport", "Country", "Total Streams", "Avg Minutes per Stream"),
      format.args = list(big.mark = ","),
      caption = "Table 3: Streaming Metrics by Sport and Country")
```
```{r sportCountryChart, fig.width=12, fig.height=6}
# Create grouped bar chart
ggplot(df_sport_country, aes(x = sport, y = total_streams, fill = country)) +
  geom_col(position = "dodge") +
  labs(
    title = "Total Streaming Events by Sport and Country",
    subtitle = "Comparison across all countries",
    x = "Sport",
    y = "Total Streaming Events",
    fill = "Country"
  ) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 0, hjust = 0.5)
  )
```

---

# Day of Week Analysis

## Peak Streaming Days by Sport and Country

This analysis identifies which day of the week has the highest streaming activity (by total minutes) over the past three years.
```{r dayOfWeekAnalysis}
# Get last 3 years
current_year <- dbGetQuery(mysqlConn, "
  SELECT MAX(dd.year) 
  FROM fact_streaming f
  JOIN dim_date dd ON f.date_id = dd.date_id
")[[1]]

three_years_ago <- current_year - 2

# Query with JOINs
sql <- paste0("
SELECT 
    ds.sport_name as sport,
    dc.country_name as country,
    dd.day_of_week,
    dd.day_name,
    SUM(f.total_minutes_streamed) as total_minutes
FROM fact_streaming f
INNER JOIN dim_date dd ON f.date_id = dd.date_id
INNER JOIN dim_sport ds ON f.sport_id = ds.sport_id
INNER JOIN dim_country dc ON f.country_id = dc.country_id
WHERE dd.year >= ", three_years_ago, "
GROUP BY ds.sport_name, dc.country_name, dd.day_of_week, dd.day_name
ORDER BY ds.sport_name, dc.country_name, total_minutes DESC
")

df_dow <- dbGetQuery(mysqlConn, sql)

# Find peak day for each sport-country combination
df_peak <- do.call(rbind, lapply(split(df_dow, list(df_dow$sport, df_dow$country)), function(x) {
  x[which.max(x$total_minutes), ]
}))

# Display table
kable(df_peak[, c("sport", "country", "day_name", "total_minutes")], 
      col.names = c("Sport", "Country", "Peak Day", "Total Minutes Streamed"),
      format.args = list(big.mark = ","),
      row.names = FALSE,
      caption = paste0("Table 4: Peak Streaming Day by Sport and Country (", three_years_ago, "-", current_year, ")"))
```
```{r dowChart, fig.width=12, fig.height=8}
# Create heatmap
ggplot(df_dow, aes(x = day_name, y = paste(sport, "-", country), fill = total_minutes)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", labels = scales::comma) +
  labs(
    title = "Streaming Activity by Day of Week",
    subtitle = paste0("Total minutes streamed (", three_years_ago, "-", current_year, ")"),
    x = "Day of Week",
    y = "Sport - Country",
    fill = "Total Minutes"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
```

---

# Conclusion
```{r conclusion}
total_streams <- format(dbGetQuery(mysqlConn, "SELECT SUM(total_streams) FROM fact_streaming")[[1]], big.mark=",")
total_minutes <- format(dbGetQuery(mysqlConn, "SELECT SUM(total_minutes_streamed) FROM fact_streaming")[[1]], big.mark=",")
avg_per_stream <- round(dbGetQuery(mysqlConn, "SELECT AVG(avg_minutes_per_stream) FROM fact_streaming")[[1]], 1)
```

This analysis of SportsTV's streaming data reveals:

- **Total Streaming Events:** `r total_streams`
- **Total Streaming Minutes:** `r total_minutes`
- **Average Minutes per Stream:** `r avg_per_stream` minutes

The data demonstrates consistent growth in streaming activity across all sports, with distinct patterns by geography and day of week. These insights can inform content acquisition strategies, marketing campaigns, and infrastructure planning.
```{r cleanup, include=FALSE}
# Disconnect from database
dbDisconnect(mysqlConn)
```

---
